---
title: "Using a VariantExperiment data object to manage COVID19 data"
author: "Jerome Ambroise, Laurent Gatto, Julie Hurel, Bertrand Bearzatto, Jean-Luc Gala"
date: '`r format(Sys.time(), "%B %d, %Y,%H:%M")`'
output:
  html_document:
    smart: FALSE
    code_folding: show
    collapsed: yes
    fig_caption: yes
    fig_height: 6
    fig_width: 9
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---

In this document we illustrate the advantages of using the
`r BiocStyle::Biocpkg("VariantExperiment")` Bioconductor
package for handling SARS-CoV-2 genomic data and associated
metadata. For more details, we invite you to read the associated
publication ('On the many advantages of using the VariantExperiment
Bioconductor package for storing, exchanging and analyzing SARS-CoV-2
genomic data and associated metadata' available on bioRxiv.org).


# Load package

```{r, message = FALSE}
library(VariantExperiment)
```

# Data loading

The `RangedSummarizedExperiment` object was created with the
VariantExperiment package for handling SARS-COV2 genomic data and
associated 'dummy' (in the present study) metadata.

```{r}
SARSCOV2 <- readRDS('SARS-CoV-2_VariantExperiment.rds')
```

# Data management

The object contains `r nrow(SARSCOV2)` variants sequenced in
`r ncol(SARSCOV2)` different samples.

```{r}
SARSCOV2
```

## `Assay`: genomic data

As shown heren, the assays slots store the Genomic data (including the
called genotype and a quality metrics such as read depth (DP) for the
illustration).  Of course other quality metrics could be added as a
phred score.

```{r}
assays(SARSCOV2)$genotype
assays(SARSCOV2)$DP
```

## `colData`: sample annotations/metadata

As shown here, the `colData` slot stores the annotation/metadata
associated to each sample (sample, pre-analytic, analytic,
bioinformatics)

```{r}
colData(SARSCOV2)
```

## `rowData`: feature annotationnns

Finally, the `rowData` includes the description of each mutation. This
slot could also include an indicator reflecting if this variant is
currently considered as a Variant of Concern (VOC) or Variant of high
Concern.

```{r}
rowData(SARSCOV2)
```


# Subsetting

Here we illustrate how to subset the object. For the demonstration we
select only data:

- from patient with obesity (Obsesity = yes)
- mutation with a start position < 5000

```{r}
COVID.FILE.Subset <- SARSCOV2[rowData(SARSCOV2)$start < 5000, SARSCOV2$Obesity == 'yes']
```

As shown here the selection is performed in one operation for both the
rowData, colData and assay(s). This property therefore ensures to keep
a perfect synchronization of genomic data with metadata.


# Session info

```{r}
sessionInfo()
```
